---
title: "Assignment 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I will start by reading the data into RStudio in two separate dataframes based on dose.

```{r Reading data in}
setwd("C:/Users/dania/Documents/Pitt/BioStats/Module 2")
x <- read.table("assignment_2_sum.txt",sep="\t",header=T)
y <- read.table("assignment_2_high.txt",sep="\t",header=T)
```
"X" is the data from patients receiving 600 mg of HCQ and "Y" is the data from patients receiving 1200 mg of HCQ. To test significant changes in PD-1 activity (defined as PD-1 MFI x cell population percentage), I will perform t tests for each cell population, comparing the low dose and high dose. To do this, I will start by entering each value under its own name.
```{r Naming cell populations}
#h corresponds to helper CD4 T cells and c corresponds to cytotoxic CD8 T cells
hmem_lo <- x$CD4mem
hmem_hi <- y$CD4mem

heff_lo <- x$CD4eff
heff_hi <- y$CD4eff

hnaive_lo <- x$CD4naive
hnaive_hi <- y$CD4naive

cmem_lo <- x$CD8mem
cmem_hi <- y$CD8mem

ceff_lo <- x$CD8eff
ceff_hi <- y$CD8eff

cnaive_lo <- x$CD8naive
cnaive_hi <- y$CD8naive
```

*Question 1*: Effect of the treatment on PD-1 activity?
Since we're using MFI x cell population % to calculate PD-1 activity, there's not much to be said about total PD-1 activity, it has to be cell type specific. I have already calculated change in PD-1 activity after treatment as % change from baseline, normalized to each patient. Now, I will run two sample t tests with Welch's correction to determine if PD-1 activity in patients with 600mg vs 1200 mg is significantly different in specific cell subsets. 

```{r T tests}
t.test(hmem_hi,hmem_lo)
t.test(heff_hi,heff_lo)
t.test(hnaive_hi,hnaive_lo)
t.test(cmem_hi,cmem_lo)
t.test(ceff_hi,ceff_lo)
t.test(cnaive_hi,cnaive_lo)
```
Noting is significant. I did not standardize among patients by accounting for differences in cell composition in samples. However, since we're technically comparing each patient to themself, I don't think this is strictly necessary. 

*Question 2*: Is PD-1 activity in different cell types independent or dependent? 
So I think we need to run a correlation matrix for each dataset (600 mg and 1200 mg) to see if PD-1 activity changes in different cell subsets is correlated with those in another subset. However, technically Chi-squared is the test of independence, so it's possible that Chi-squared is the only test for true independence.
```{r Correlation matrices}
#I started by trying to just run cor() on x and y but it didn't work, so I made matrices of each dataset with just the % change from baseline values (no headings, no patient numbers, no dosage) so that all of the values would be numeric
setwd("C:/Users/dania/Documents/Pitt/BioStats/Module 2")
q <- read.table("assignment_2_high_matrix.txt",sep="\t",header=FALSE)
q
cor(q,method = "pearson") #1200 dose, looking at correlations between pairs of cell types
plot(q)

v <- read.table("assignment_2_sum_matrix.txt",sep="\t",header=FALSE)
v
cor(v,method = "pearson") #600 dose, looking at correlations between pairs of cell types
plot(v)

#now I'm wondering if I need to do a Chi-squared test, but using the chisq.test() function for q and v did not work because the values have to be "nonnegative and finite"
a <- abs(q) #taking the absolute value of q and putting it into a new variable
#now I can run the Chi-squared test
chisq.test(a)
#I don't think this preserves the data accurately 
```

