---
title: "Assignment 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I will start by reading the data into RStudio in two separate dataframes based on dose.

```{r Reading data in}
setwd("C:/Users/dania/Documents/Pitt/BioStats/Module 2")
x <- read.table("assignment_2_sum.txt",sep="\t",header=T)
y <- read.table("assignment_2_high.txt",sep="\t",header=T)
```
"X" is the data from patients receiving 600 mg of HCQ and "Y" is the data from patients receiving 1200 mg of HCQ. To test significant changes in PD-1 activity (defined as PD-1 MFI x cell population percentage), I will perform t tests for each cell population, comparing the low dose and high dose. To do this, I will start by entering each value under its own name.
```{r Naming cell populations}
#h corresponds to helper CD4 T cells and c corresponds to cytotoxic CD8 T cells
hmem_lo <- x$CD4mem
hmem_hi <- y$CD4mem

heff_lo <- x$CD4eff
heff_hi <- y$CD4eff

hnaive_lo <- x$CD4naive
hnaive_hi <- y$CD4naive

cmem_lo <- x$CD8mem
cmem_hi <- y$CD8mem

ceff_lo <- x$CD8eff
ceff_hi <- y$CD8eff

cnaive_lo <- x$CD8naive
cnaive_hi <- y$CD8naive
```

```{r T tests}
t.test(hmem_hi,hmem_lo)
t.test(heff_hi,heff_lo)
t.test(hnaive_hi,hnaive_lo)
t.test(cmem_hi,cmem_lo)
t.test(ceff_hi,ceff_lo)
t.test(cnaive_hi,cnaive_lo)
```
```{r}
cor(x,y, method = "pearson") #this is not working and I'm not entirely sure why. might try to run a linear regression
```

