---
title: "Assignment 3"
author: "Dani Antos"
date: "7/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3

Going to use a linear regression model for each cell population as the dependent variable. It's going to be a multiple regression with interactions, but there doesn't need to be quadratics involved (I don't think)

Y = B0 + B1(ATP6) + B2(COX1) + B3(CYTB) + B4(ND4) + B5(ND5) + B6(ATP6xCOX1) + B7(ATP6xCYTB) + B8(COX1xCYTB) + B9(COX1xND5) + B10(CYTBxND5) + B11(ND4xND5) + B12(ATP6xCOX1xCYTBxND4xND5)

```{r}
setwd("C:/Users/dania/Documents/Pitt/BioStats/Module 3")
z <- read.table("t_cells.txt",sep="\t",header=T)
CD.3 <- lm(CD3.pct ~ ATP6 + COX1 + CYTB + ND4 + ND5 + ATP6*COX1 + ATP6*CYTB + COX1*CYTB + COX1*ND5 + CYTB*ND5 + ND4*ND5 + ATP6*COX1*CYTB*ND4*ND5, data=z, na.rm=TRUE)

anova(CD.3)
library(car)
Anova(CD.3, type=2)
qqnorm(CD.3$residuals)
qqline(CD.3$residuals)

CD.8 <- lm(CD8.pct ~ ATP6 + COX1 + CYTB + ND4 + ND5 + ATP6*COX1 + ATP6*CYTB + COX1*CYTB + COX1*ND5 + CYTB*ND5 + ND4*ND5 + ATP6*COX1*CYTB*ND4*ND5, data=z, na.rm=TRUE)
summary(CD.8)

anova(CD.8)
library(car)
Anova(CD.8, type=2)
qqnorm(CD.8$residuals)
qqline(CD.8$residuals)

FOXP3 <- lm(FOXP3.pct ~ ATP6 + COX1 + CYTB + ND4 + ND5 + ATP6*COX1 + ATP6*CYTB + COX1*CYTB + COX1*ND5 + CYTB*ND5 + ND4*ND5 + ATP6*COX1*CYTB*ND4*ND5, data=z, na.rm=TRUE)
summary(FOXP3)

anova(FOXP3)
library(car)
Anova(FOXP3, type=2)
qqnorm(FOXP3$residuals)
qqline(FOXP3$residuals)

CD31 <- lm(CD31.pct ~ ATP6 + COX1 + CYTB + ND4 + ND5 + ATP6*COX1 + ATP6*CYTB + COX1*CYTB + COX1*ND5 + CYTB*ND5 + ND4*ND5 + ATP6*COX1*CYTB*ND4*ND5, data=z, na.rm=TRUE)
summary(CD31)

anova(CD31)
library(car)
Anova(CD31, type=2)
qqnorm(CD31$residuals)
qqline(CD31$residuals)

CD31.2 <- lm(CD31.pct ~ ATP6 + ND5, data=z, na.rm=TRUE)
summary(CD31.2)
anova(CD31.2)
library(car)
Anova(CD31.2, type=2)
qqnorm(CD31.2$residuals)
qqline(CD31.2$residuals)

CD31.1 <- lm(CD31.pct ~ COX1 + ND5, data=z, na.rm=TRUE)
summary(CD31.1)

anova(CD31.1)
library(car)
Anova(CD31.1, type=2)
qqnorm(CD31.1$residuals)
qqline(CD31.1$residuals)
```

### Potential Transformation Options for Numeric Data

library(lme4)

q <- read.table("t_cells_1.txt",sep="\t",header=T)

lmer(staining ~ ATP6 + COX1 + CYTB + ND4 + ND5 + ATP6xCOX1 + ATP6xCYTB + COX1xCYTB + COX1xND5 + CYTBxND5 + ND4xND5 + ATP6xCOX1xCYTBxND4xND5 + (1|subject), data=q)

CD.3.1 <- lm(CD3.pct ~ ATP6 + COX1 + CYTB + ND4 + ND5 + ATP6^2 + COX1^2 + CYTB^2 + ND4^2 + ND5^2 + ATP6xCOX1 + ATP6xCYTB + COX1xCYTB + COX1xND5 + CYTBxND5 + ND4xND5 + ATP6xCOX1xCYTBxND4xND5, data=z, na.rm=TRUE)

summary(CD.3.1)

library(gam)

gam <- gam(CD3.pct ~ ATP6 + COX1 + CYTB + ND4 + ND5 + ATP6xCOX1 + ATP6xCYTB + COX1xCYTB + COX1xND5 + CYTBxND5 + ND4xND5 + ATP6xCOX1xCYTBxND4xND5, data = z)

gam

summary(gam) #this is EXACTLY the same as the anova table (type 1) using the linear model in CD.3

library(effects)

plot(effects(CD.3))

crPlots(CD.3) is in the car package to find nonlinear relations between independent and dependent variables but you can't use it with models that have interactions.

boxTidwell(CD3.pct ~ ATP6 + COX1 + CYTB + ND4 + ND5 + ATP6^2 + COX1^2 + CYTB^2 + ND4^2 + ND5^2 + ATP6xCOX1xCYTBxND4xND5, data=z) #values have to be positive